---
title: "FMCG Promotion Effectiveness Analysis"
author: "Jirawan Phromtawepong"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(dplyr)
library(ggplot2)
library(scales)
library(lubridate)
```

# 📦 Dataset Overview

```{r load-data}
Superstore <- read_csv("Sample - Superstore.csv")

# Fix column names if needed
glimpse(Superstore)
```

# 🧼 Data Cleaning

```{r clean-data}
# Convert Order Date to proper date
Superstore$`Order Date` <- mdy(Superstore$`Order Date`)

# Drop unnecessary columns (optional)
Superstore <- Superstore %>%
  select(-`Row ID`, -`Postal Code`)

# Confirm structure
glimpse(Superstore)
```

# 📊 Sub-Category Performance Summary

```{r summary-subcat}
summary_by_subcat <- Superstore %>%
  group_by(`Sub-Category`) %>%
  summarise(
    Total_Sales = sum(Sales),
    Total_Profit = sum(Profit),
    Avg_Discount = mean(Discount),
    Profit_Margin = sum(Profit) / sum(Sales)
  ) %>%
  arrange(desc(Total_Sales))

summary_by_subcat
```

# 📈 Profit Margin by Sub-Category

```{r profit-margin-barplot}
ggplot(summary_by_subcat, aes(x = reorder(`Sub-Category`, Profit_Margin), y = Profit_Margin)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Profit Margin by Sub-Category",
       x = "Sub-Category",
       y = "Profit Margin") +
  scale_y_continuous(labels = percent_format(accuracy = 1))
```

# 📉 Discount vs Profit

```{r discount-vs-profit}
ggplot(Superstore, aes(x = Discount, y = Profit)) +
  geom_point(alpha = 0.4, color = "tomato") +
  geom_smooth(method = "lm", se = FALSE, color = "steelblue") +
  labs(title = "Discount vs Profit",
       x = "Discount",
       y = "Profit")
```

# 🔍 Insight Summary

- Sub-categories like **Copiers** and **Accessories** have strong margins, while **Tables** show negative profitability.
- **Deep discounts (≥ 40%)** often lead to **loss-making transactions**.
- Many profitable transactions occur at **0–10% discounts**, suggesting opportunities to scale back promotions on resilient items.
